animbook: Visualizing changes in performance measures and demographic affiliations using animation

### 2024-01-18 animbook-journal.Rmd
polishing conclusions and usage example

In this section, the package usage will be demonstrated using the `cat_change` dataset included in this package. This dataset is simulated data that has changed from category A to E between two time points. Since this data is already in the categorized data structure, the `anim_prep_cat()` function is utilized. The output produced from this package is referred to as the `categorized` data.

```{r, echo=TRUE}
animbook <- anim_prep_cat(cat_change, 
                          id = id, 
                          values = qnt, 
                          time = time, 
                          group = gp)

head(animbook, 10)
str(animbook)
```

  To obtain a `ggplot` object for the animation function, the `wallaby_plot` is used. For further customization, the `shade_palette`, `subset`, and `relation` arguments are applied to define the final appearance of the animation.

```{r, echo=TRUE, fig.width=8, fig.align='center', out.width="100%", layout = "l-body"}
p <- wallaby_plot(data = animbook,
             group_palette = RColorBrewer::brewer.pal(9, "Set1"),
             shade_palette = c("#737373", "#969696", "#BDBDBD",
                               "#D9D9D9","#D9D9D9","#D9D9D9"),
             rendering = "gganimate",
             subset = "top",
             relation = "one_many",
             total_point = NULL)

@@

  Once the `ggplot` object is obtained, the `anim_animate()` function can be used to transform it into a `gganimate` object. This then can be passed onto the `animate()` function from the \CRANpkg{gganimate} for rendering.

@@

  This is an example of how the users can apply the three-step process in creating an animation plot using the `animbook` package.

@@

Beginning with inspiration drawn from the New York Times article, this packages provides tools to facilitate the communication of complex data to broader audiences. The paper outlines the anticipated data format for input into the prep function, covering raw, categorized, and animated data formats. It delves into the animation concept, persistence of vision, achieved through point interpolation, and the visualization design choices.

  The \CRANpkg{animbook} package comes with certain drawbacks. Firstly, it leans more towards being a communication tool rather than an exploratory one, as the information it presents can often be obtained from a summary table. Secondly, it may struggle to handle datasets with numerous categories (quantiles), limiting its applicability in such scenarios. Furthermore, perceiving patterns from the animation might be challenging for users when there are differences in sample sizes across groups. Lastly, the animation requires the reader to mentally calculate the fraction of points that are in motion. However, the package provides tools that provides a simple solution for creating meaningful animations through a simple step process.

  Introducing features that allow variation in speeds for each observation could enhance the visualization by introducing an additonal dimension. The new animated and static plots, as demonstrated in the New York Times article, would enable comparisons of demographics at specific points in time. Additionally, the potential re-parameterization of the x-axis from years to ages (e.g., company ages) could add different ways of viewing the visualization.

- the first paragraph is to summarize the achievement of this
- the second paragraph is about the pro and con, what works, what doesn't works.
- the third paragraph, if you have another 6 months, what would you implemented.

>>>>

The `cat_change` data in the package is used to illustrate how to use this package. It has five categories, A to E, over two time points. Since the data is already in the categorized data structure, the `anim_prep_cat()` function is used. The output from this function is a class called `categorized`.

```{r, echo=TRUE}
a <- anim_prep_cat(cat_change, 
                          id = id, 
                          values = qnt, 
                          time = time, 
                          group = gp)
class(a)
glimpse(a)
```

  To obtain an object for the animation function, the `wallaby_plot` is used. The `shade_palette`, `subset`, and `relation` arguments are applied to refine the final appearance of the animation. REFER TO FIGURE

```{r, echo=TRUE, fig.width=8, fig.align='center', out.width="100%", layout = "l-body", fig.cap="FIXME"}
p <- wallaby_plot(data = a,
             group_palette = RColorBrewer::brewer.pal(9, "Set1"),
             rendering = "gganimate",
             subset = "top",
             relation = "one_many",
             total_point = NULL)

@@

  Users have the flexibility to use various options, such as `subset` or `relation`, to customize the plot appearance. Here "top" is used to to specify category A for the final time point, and "many_one" is used to specify that all categories flowing into A at time point 2 are shown. REFER TO FIGURE

```{r fig.cap="FIXME"}
p_2 <- wallaby_plot(data = a,
                   rendering = "gganimate",
                   subset = "top",
                   relation = "many_one",
                   total_point = NULL)

p_2
```

Using `subset = "B"` will only show the point for a single category. REFER TO FIGURE

```{r fig.cap="FIXME"}
p_3 <- wallaby_plot(data = a,
                    rendering = "gganimate",
                    subset = "B",
                    relation = "many_one",
                    total_point = NULL)

p_3
```

  This plot object is passed to the `anim_animate()` function to transform it into a `gganimate` object, which appears as an animation when the `animate()` function from \CRANpkg{gganimate} is used for rendering. REFER TO PLOT

@@

@@

Beginning with inspiration drawn from the New York Times article, this package provides tools to generate animations for temporal flow data to study changes in groups over time. The paper describes how to pre-process the data into a format for making several types of plots and animation. It also explains what needs to be considered when making an animation such as persistence of objects, how to achieve this using interpolation, and the visualization design choices automated and those available to users.

  The package provides the tools to create animations for simple temporal patterns, with a few steps, but these visualization methods have some limitations. It is primarily for communication rather than exploration because one typically knows what the main patterns are. If there are many categories (quantiles) the display will be too crowded. Because the animation involves perceiving patterns in groups of moving points it can be challenging with unbalanced sample sizes. This requires the reader to mentally calculate the fraction of points that are in motion.

  There are some potential new directions to develop this work. Some applications might benefit by re-parameterization of the x-axis from years to ages (e.g. company ages, time since the company was founded). One could also re-arrange the mapping of variables to the plot and animations. For example, the New York Times article, mentioned that to better compare distributions in time, one could plot the quantiles, which would form something like an S-curve, and animate over time, to observe how the distributions are changing.
